### 2024-04-05 doc/15yrsChange.Rmd
discussion
- new line per sentence (easier for editing)
- a few changes in first couple sentences
- commented out future directions

The combination of -∆AGB and limited recruitment of canopy species in the understory is compromising the potential carbon sink of this forest, now and into the future. Decreases in AGB were largest in the regions of the plot with high canopy vulnerability to pests and pathogens. These regions also exhibited the lowest recruitment of canopy species in the understory. Deer severely depleted regeneration of canopy species and the region with low deer density maintained the highest amount of canopy species in the understory. Despite localized increases in AGB, notably in areas with low canopy vulnerability, on net, these increases are not outpacing AGB lost through a combination of increased AWM and decreasing AWG. Total ABG stocks in the plot were lower in 2023 than in 2013. Moreover, the low recruitment of canopy species in areas with high canopy vulnerability suggests that heavy deer browsing on canopy species may be exacerbated in areas with canopy openings due to nuisance-species driven mortality where fast-growing understory species that are resistant to deer can take advantage, further competing with canopy recruitment.

@@

# Future Directions

@@

Future research should build on our results by expanding these analyses to other temperate forest regions. While other areas will have different forest compositions and herbivory dynamics, increasing pressure from pests and pathogens is pervasive throughout temperate forests (Santini et al. 2012, Ayres and Lombardero 2018). Moreover, while not a large factor at this site, the spread of non-indigenous plant species can further threaten the  regeneration of woody canopy species (Pile Knapp et al, 2023). Large forest dynamics sites (*i.e.* other sites in the ForestGEO network) are well-suited to this research and can be leveraged to quantify how nuisance species compromise present and future forest carbon sequestration. Expanding the scale of these analyses may also facilitate the incorporation of nuisance species into global vegetation and carbon models. Future evaluations of temperate forest health and carbon storage must account for the substantial effects of nuisance species along with other global changes.

>>>>

Our results reveal that aThe combination of declining biomass (-∆AGB) and limited recruitment of canopy species in the understory is compromising the potential carbon sink of the SCBI ForestGEO plot, now and into the future. 
Decreases in AGB were largest in the plot areas with high canopy vulnerability to pests and pathogens.
These regions also exhibited the lowest recruitment of canopy species in the understory. Deer severely depleted regeneration of canopy species and the region with low deer density maintained the highest amount of canopy species in the understory.
Despite localized increases in AGB, notably in areas with low canopy vulnerability, on net, these increases are not outpacing AGB lost through a combination of increased AWM and decreasing AWG.
Total ABG stocks in the plot were lower in 2023 than in 2013. Moreover, the low recruitment of canopy species in areas with high canopy vulnerability suggests that heavy deer browsing on canopy species may be exacerbated in areas with canopy openings due to nuisance-species driven mortality where fast-growing understory species that are resistant to deer can take advantage, further competing with canopy recruitment.

@@

<!--# Future Directions

@@

Future research should build on our results by expanding these analyses to other temperate forest regions. While other areas will have different forest compositions and herbivory dynamics, increasing pressure from pests and pathogens is pervasive throughout temperate forests (Santini et al. 2012, Ayres and Lombardero 2018). Moreover, while not a large factor at this site, the spread of non-indigenous plant species can further threaten the  regeneration of woody canopy species (Pile Knapp et al, 2023). Large forest dynamics sites (*i.e.* other sites in the ForestGEO network) are well-suited to this research and can be leveraged to quantify how nuisance species compromise present and future forest carbon sequestration. Expanding the scale of these analyses may also facilitate the incorporation of nuisance species into global vegetation and carbon models. Future evaluations of temperate forest health and carbon storage must account for the substantial effects of nuisance species along with other global changes. -->
